---
title: "Topic modeling - Latent Semantic Analysis(LSA) "
output: 
  rmarkdown::html_vignette:
    css: mystyle.css
    toc: yes
vignette: >
  %\VignetteIndexEntry{Text Analysis for Literature}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Quickstart}
-->

In this vignette we show how to perform Latent Semantic Analysis using the **quanteda** package. Example is taken from Grossman and Frieder’s Information Retrieval, Algorithms and Heuristics. 

# 1 Create a document-feature matrix

```{r}
require(quanteda, quietly = TRUE, warn.conflicts = FALSE)
txt <- c(d1="Shipment of gold damaged in a fire",
d2="Delivery of silver arrived in a silver truck",
d3="Shipment of gold arrived in a truck" )

mydfm <- dfm(txt)
mydfm
```



# 2. Construct a lsa model
```{r}
mylsa <- textmodel_lsa(mydfm)
```

 the new document vector coordinates in the reduced 2-dimensional space is:
```{r}
mylsa$docs[, 1:2]
```

# 3. Apply the construted lsa model to new data
Now the new unseen document can be represented in the reduced 2-dimensional space.
The unseen query document:
```{r}
query <- dfm(c("gold silver truck"))
cdfm <- rbind(mydfm, query)
querydfm <- cdfm[nrow(cdfm), ]
querydfm
```
```{r}
newq <- predict(mylsa, querydfm)
newq$transDocs[, 1:2]
```
