---
title: "Topic modeling - Latent Semantic Analysis (LSA)"
author: Haiyan Wang
output: 
  rmarkdown::html_vignette:
    css: mystyle.css
    toc: yes
vignette: >
  %\VignetteIndexEntry{Latent Semantic Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Quickstart}
-->

In this vignette we show how to perform Latent Semantic Analysis using the **quanteda** package. Example is taken from Grossman and Frieder’s _Information Retrieval, Algorithms and Heuristics_, based on the worked example (here)[http://www1.se.cuhk.edu.hk/%7Eseem5680/lecture/LSI-Eg.pdf].

LSA decomposes document-feature matrix into a reduced vector space that is assumed to reflect semantic structure.

New documents or queries can be 'folded-in' to this constructed latent semantic space for downstream tasks. 

## 1. Create a document-feature matrix

```{r}
require(quanteda, quietly = TRUE, warn.conflicts = FALSE)
txt <- c(d1="Shipment of gold damaged in a fire",
d2="Delivery of silver arrived in a silver truck",
d3="Shipment of gold arrived in a truck" )

mydfm <- dfm(txt)
mydfm
```



## 2. Construct the LSA model
```{r}
mylsa <- textmodel_lsa(mydfm)
```

 the new document vector coordinates in the reduced 2-dimensional space is:
```{r}
mylsa$docs[, 1:2]
```

## 3. Apply the constructed LSA model to new data
Now the new unseen document can be represented in the reduced 2-dimensional space.
The unseen query document:
```{r}
querydfm <- dfm(c("gold silver truck")) %>%
    dfm_select(pattern = mydfm)
querydfm
```
```{r}
newq <- predict(mylsa, querydfm)
newq$docs_newspace[, 1:2]
```
