---
title: "Quanteda"
output: 
  rmarkdown::html_vignette:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "man/images/"
)
```

```{r, echo=FALSE}
knitr::include_graphics("https://cdn.rawgit.com/kbenoit/quanteda/master/images/quanteda_logo.svg")
```


## About

An R package for managing and analyzing text, created by [Kenneth Benoit](http://kenbenoit.net) in 
collaboration with a team of core 
[contributors](https://github.com/kbenoit/quanteda/graphs/contributors): 
[Kohei Watanabe](https://github.com/koheiw), 
[Paul Nulty](https://github.com/pnulty), 
[Adam Obeng](https://github.com/adamobeng), 
[Haiyan Wang](https://github.com/HaiyanLW), 
[Ben Lauderdale](https://github.com/lauderdale), 
and [Will Lowe](https://github.com/conjugateprior).  
Supported by the European Research Council grant ERC-2011-StG 283794-QUANTESS.

For more details, see http://quanteda.io.

How to cite the package:

```{r echo = FALSE, out.width = 80, comment = NA}
citation("quanteda")
```

## How to Install

1.  From [CRAN](https://CRAN.R-project.org/package=quanteda): Use your GUI's R package installer, or execute:

    ```{r eval = FALSE}
    install.packages("quanteda") 
    ```

2.  From [GitHub](https://github.com/kbenoit/quanteda), using:

    ```{r eval = FALSE}
    # devtools packaged required to install quanteda from Github 
    devtools::install_github("kbenoit/quanteda") 
    ```

    Because this compiles some C++ source code, you will need a compiler installed. 
    If you are using a Windows platform, this means you will need also to install 
    the [Rtools](https://CRAN.R-project.org/bin/windows/Rtools/) software available 
    from CRAN.  **If you are using macOS**, you will need to to install XCode, available 
    for free from the App Store, or if you prefer a lighter footprint set of tools, 
    [just the Xcode command line 
    tools](http://osxdaily.com/2014/02/12/install-command-line-tools-mac-os-x/ ), 
    using the command `xcode-select --install` from the Terminal.
    
    Also, you might need to upgrade your compiler.  @kbenoit found that his macOS build only worked reliably after upgrading the default Xcode compiler to clang4, following [these instructions](http://thecoatlessprofessor.com/programming/openmp-in-r-on-os-x/#after-3-4-0).


## Leaving feedback

If you like **quanteda**, please consider leaving [feedback or a testimonial here](https://github.com/kbenoit/quanteda/issues/461).



## Contributing

Contributions in the form of feedback, comments, code, and bug reports are most welcome.  How to contribute:

*  Fork the source code, modify, and issue a [pull request](https://help.github.com/articles/creating-a-pull-request-from-a-fork/) through the [project GitHub page](https://github.com/kbenoit/quanteda).  See our [Contributor Code of Conduct](https://github.com/kbenoit/quanteda/blob/master/CONDUCT.md) and the all-important **quanteda** [Style Guide](https://github.com/kbenoit/quanteda/wiki/Style-guide).
*  Issues, bug reports, and wish lists: [File a GitHub issue](https://github.com/kbenoit/quanteda/issues).
*  Usage questions:  Submit a question on the [**quanteda** channel on StackOverflow](http://stackoverflow.com/questions/tagged/quanteda).
*  Contact [the maintainer](kbenoit@lse.ac.uk) by email.

