---
output: github_document
---


<!-- README.md is generated from README.Rmd. Please edit that file -->
```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "man/images/"
)
```
###[![quanteda: quantitative analysis of textual data](https://cdn.rawgit.com/kbenoit/quanteda/master/images/quanteda_logo.svg)](http://quanteda.io)

[![CRAN Version](https://www.r-pkg.org/badges/version/quanteda)](https://CRAN.R-project.org/package=quanteda)
[![Downloads](https://cranlogs.r-pkg.org/badges/quanteda)](https://CRAN.R-project.org/package=quanteda)
[![Total Downloads](https://cranlogs.r-pkg.org/badges/grand-total/quanteda?color=orange)](https://CRAN.R-project.org/package=quanteda)
[![Travis-CI Build Status](https://travis-ci.org/kbenoit/quanteda.svg?branch=master)](https://travis-ci.org/kbenoit/quanteda) 
[![Build status](https://ci.appveyor.com/api/projects/status/e3tf2h1ff0nlv249/branch/master?svg=true)](https://ci.appveyor.com/project/kbenoit/quanteda/branch/master)
[![codecov.io](https://codecov.io/github/kbenoit/quanteda/coverage.svg?branch=master)][1]

[1]: https://codecov.io/gh/kbenoit/quanteda/branch/master


## **quanteda** v0.99 Important Changes

Many important changes to the package have been underway, including API changes, as we approach a stable API, "1.0" release in October 2017.  Version 0.99 represents the last version that will contain many of the deprecated object types and methods that date several releases.   

v0.99 also implements many enhancements and performance improvements over previous releases.  See [NEWS.md](http://quanteda.io/news/index.html#quanteda-0-99) for details, and [Quanteda Structure and Design](https://kbenoit.github.io/quanteda/articles/development-plans.html) for a description of the package's underlying logic and design philosophy.


## About the package

An R package for managing and analyzing text, created by [Kenneth Benoit](http://kenbenoit.net) in 
collaboration with a team of core 
[contributors](https://github.com/kbenoit/quanteda/graphs/contributors): 
[Kohei Watanabe](https://github.com/koheiw), 
[Paul Nulty](https://github.com/pnulty), 
[Adam Obeng](https://github.com/adamobeng), 
[Haiyan Wang](https://github.com/HaiyanLW), 
[Ben Lauderdale](https://github.com/lauderdale), 
and [Will Lowe](https://github.com/conjugateprior).  
Supported by the European Research Council grant ERC-2011-StG 
283794-QUANTESS.

For more details, see the [**package website**](http://quanteda.io).

How to cite the package:

```{r echo = FALSE, comment = NA}
citation("quanteda")
```


### Leave feedback

If you like **quanteda**, please consider leaving [feedback or a testimonial here](https://github.com/kbenoit/quanteda/issues/461).


## How to Install

1.  From [CRAN](https://CRAN.R-project.org/package=quanteda): Use your GUI's R package installer, or execute:

    ```{r eval = FALSE}
    install.packages("quanteda") 
    ```

2.  From [GitHub](https://github.com/kbenoit/quanteda), using:

    ```{r eval = FALSE}
    # devtools packaged required to install quanteda from Github 
    devtools::install_github("kbenoit/quanteda") 
    ```

    Because this compiles some C++ source code, you will need a compiler installed. 
    If you are using a Windows platform, this means you will need also to install 
    the [Rtools](https://CRAN.R-project.org/bin/windows/Rtools/) software available 
    from CRAN.  If you are using OS X, you will need to to install XCode, available 
    for free from the App Store, or if you prefer a lighter footprint set of tools, 
    [just the Xcode command line 
    tools](http://osxdaily.com/2014/02/12/install-command-line-tools-mac-os-x/ ), 
    using the command `xcode-select --install` from the Terminal.

3.  Additional recommended packages:

    The following packages work well with or extend **quanteda** and we recommend that you also install them:
    
    *  [**readtext**](https://github.com/kbenoit/readtext):  An easy way to read text data into R, from almost any input format.
    
    *  [**spacyr**](https://github.com/kbenoit/spacyr): NLP using the [spaCy](http://spacy.io) library, including part-of-speech tagging, entity recognition, and dependency parsing.

    *  [**quantedaData**](https://github.com/kbenoit/quantedaData): Additional textual data for use with **quanteda**.

        ```{r eval = FALSE}
        devtools::install_github("kbenoit/quantedaData")
        ```
    *  [**LIWCalike**](https://github.com/kbenoit/LIWCalike): An R implementation of the [Linguistic Inquiry and Word Count](http://liwc.wpengine.com) approach to text analysis.
    
        ```{r eval = FALSE}
        devtools::install_github("kbenoit/LIWCalike")
        ```
    
        
## Getting Started

See the [package website](http://quanteda.io), which includes the [Getting Started 
Vignette](http://quanteda.io/articles/quickstart.html).


## Demonstration

```{r quanteda_example, fig.width=7, fig.height=7}
library(quanteda)

# create a corpus from the immigration texts from UK party platforms
uk2010immigCorpus <- 
    corpus(data_char_ukimmig2010,
           docvars = data.frame(party = names(data_char_ukimmig2010)),
           metacorpus = list(notes = "Immigration-related sections of 2010 UK party manifestos"))
uk2010immigCorpus
summary(uk2010immigCorpus)

# key words in context for "deport", 3 words of context
kwic(uk2010immigCorpus, "deport", 3)

# create a dfm, removing stopwords
mydfm <- dfm(uk2010immigCorpus, remove = stopwords("english"), remove_punct = TRUE)
mydfm

topfeatures(mydfm, 20)  # 20 top words

# plot a word cloud
set.seed(100)
textplot_wordcloud(mydfm, min.freq = 6, random.order = FALSE,
                   rot.per = .25, 
                   colors = RColorBrewer::brewer.pal(8,"Dark2"))
```


## Contributing

Contributions in the form of feedback, comments, code, and bug reports are most welcome.  How to contribute:

*  Fork the source code, modify, and issue a [pull request](https://help.github.com/articles/creating-a-pull-request-from-a-fork/) through the [project GitHub page](https://github.com/kbenoit/quanteda).  See our [Contributor Code of Conduct](https://github.com/kbenoit/quanteda/blob/master/CONDUCT.md) and the all-important **quanteda** [Style Guide](https://github.com/kbenoit/quanteda/wiki/Style-guide).

*  Issues, bug reports, and wish lists: [File a GitHub issue](https://github.com/kbenoit/quanteda/issues).

*  Usage questions:  Submit a question on the [**quanteda** channel on StackOverflow](http://stackoverflow.com/questions/tagged/quanteda).

*  Contact [the maintainer](kbenoit@lse.ac.uk) by email.

