---
title: "Comparison to other packages"
author: Stefan Müller, Kenneth Benoit, and Kohei Watanabe
output: 
  html_document:
    toc: true
---

```{r eval = TRUE, message = FALSE, warning = FALSE, echo = FALSE}
library(quanteda)
library(knitr)
```
This article compares **quanteda** to alternative R packages for quantitative text analysis ([**tm**](https://cran.r-project.org/web/packages/tm/index.html), [**tidytext**](https://cran.r-project.org/web/packages/tidytext/index.html), [**corpus**](https://cran.r-project.org/web/packages/corpus/index.html), and [**koRpus**](https://cran.r-project.org/web/packages/koRpus/)) and the [Natural Language Toolkit](http://www.nltk.org/index.html) for Python. If a function is available in another package, we provide the respective command. 

Note that we have used the package manuals for the comparison. If we have overlooked certain functions, please let us know - either by editing the table and issuing a Pull Request or by contacting the [maintainer](K.R.Benoit@lse.ac.uk).

```{r eval = TRUE, message = FALSE, echo = FALSE, results = 'asis'}
dta_table <- read.csv("comparison-packages-data.csv")
dta_table <- dta_table[-2] # exclude "Category" to save space
kable(dta_table, format = "markdown")
```


