---
title: "Use Case 1: Project Gutenberg Texts"
output:
  rmarkdown::html_document:
    theme: null
    css: mystyle.css
    toc: yes
vignette: >
  %\VignetteIndexEntry{Quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Quickstart}
-->

# Example 1: Moby Dick

In this use case I will implement some of the analysis from the book "Text Analysis with R for Students of Literature" by Matthew L Jockers. The first section of analysis uses a 

## scanning text and lines from the web

The code below scans and splits the text of Moby Dick from Project Gutenberg, more or less as implemented in the textbook.
```{r eval=TRUE}
library(quanteda)
alltext <- scan("http://www.gutenberg.org/cache/epub/2701/pg2701.txt",
what="character", sep="\n")
mobyWords <- unlist(strsplit(alltext, "\\W"))
which(mobyWords=="whale")
```

Below is the initial word frequency analysis performed using a `quanteda` corpus.

```{r eval=TRUE}
mobyLines <- scan("http://www.gutenberg.org/cache/epub/2701/pg2701.txt",
                what="character", sep="\n")
mobyText <- paste(mobyLines, collapse=" ")
mobyCorp <- corpus(mobyText)
```


## Comparing word frequencies

```{r eval=TRUE}
# ten most frequent words
mobyDfm <- dfm(mobyCorp)
topfeatures(mobyDfm, n=10)
plot(topfeatures(mobyDfm, n=10))

# frequencies of 'he' and 'she' - these are matrixes, not numerics
mobyDfm[,'he']
mobyDfm[,'she']


# relative frequencies:
mobyDfm <- weight(mobyDfm, type='relFreq')
mobyDfm[,'he']
mobyDfm[,'she']
```


## Dispersion plots

Example 4.1 in the book plots the frequency of certain words across the length of the text. 

```{r eval=TRUE}
nTime <- seq(1:length(mobyWords))
whales <- which(mobyWords == "whale")
wcount <- rep(NA,length(nTime))
wcount[whales] <- 1
plot(wcount, main="Dispersion Plot of `whale' in Moby Dick",
xlab="Novel Time", ylab="whale", type="h", ylim=c(0,1), yaxt='n')
```


## Searching and breaking by chapter
```{r eval=TRUE}
txt <- paste(scan(url("http://www.gutenberg.org/cache/epub/2701/pg2701.txt"), what="character", sep="\n"), collapse="\n")
#kwic(mobyCorp, 'chapter')
#chapters <- segment(mobyCorp, delimiter="Chapter")
```





