---
title: "Quick Start Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The Structure of quanteda}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = FALSE, 
                      comment = "##")
```

# Installing the package

Since **quanteda** is available on [CRAN](https://CRAN.R-project.org/package=quanteda), you can install  by using your GUI's R package installer, or execute:

```{r, eval = FALSE}
install.packages("quanteda") 
```
See an instructions at https://github.com/quanteda/quanteda to install the GitHub version, 

## Additional recommended packages:

The following packages work well with or extend **quanteda** and we recommend that you also install them:

*  [**readtext**](https://github.com/quanteda/readtext):  An easy way to read text data into R, from almost any input format.
*  [**spacyr**](https://github.com/quanteda/spacyr): NLP using the [spaCy](http://spacy.io) library, including part-of-speech tagging, entity recognition, and dependency parsing.
*  [**quanteda.corpora**](https://github.com/quanteda/quanteda.corpora): Additional textual data for use with **quanteda**.
    ```{r eval = FALSE}
    devtools::install_github("quanteda/quanteda.corpora")
    ```
*  [**quanteda.dictionaries**](https://github.com/kbenoit/quanteda.dictionaries): Various dictionaries for use with **quanteda**, including the function `liwcalike()`, an R implementation of the [Linguistic Inquiry and Word Count](http://liwc.wpengine.com) approach to text analysis.
    ```{r eval = FALSE}
    devtools::install_github("kbenoit/quanteda.dictionaries")
    ```

# Creating a Corpus

You load the package to access to functions and data in the package.

```{r, message = FALSE}
library(quanteda)
```

## Currently available corpus sources

**quanteda** has a simple and powerful companion package for loading texts: [**readtext**](https://github.com/quanteda/readtext). The main function in this package, `readtext()`, takes a file or fileset from disk or a URL, and returns a type of data.frame that can be used directly with the `corpus()` constructor function, to create a **quanteda** corpus object.

`readtext()` works on:

* text (`.txt`) files;
* comma-separated-value (`.csv`) files;
* XML formatted data;
* data from the Facebook API, in JSON format;
* data from the Twitter API, in JSON format; and
* generic JSON data.

The corpus constructor command `corpus()` works directly on:

* a vector of character objects, for instance that you have already loaded into the workspace using other tools;
* a `VCorpus` corpus object from the **tm** package.
* a data.frame containing a text column and any other document-level metadata.

### Building a corpus from a character vector

The simplest case is to create a corpus from a vector of texts already in memory in R. This gives the advanced R user complete flexbility with his or her choice of text inputs, as there are almost endless ways to get a vector of texts into R.

If we already have the texts in this form, we can call the corpus constructor function directly. We can demonstrate this on the built-in character object of the texts about immigration policy extracted from the 2010 election manifestos of the UK political parties (called `data_char_ukimmig2010`).

```{r}
my_corpus <- corpus(data_char_ukimmig2010)  # build a new corpus from the texts
summary(my_corpus)
```

If we wanted, we could add some document-level variables -- what quanteda calls *docvars* -- to this corpus.

We can do this using the R's `names()` function to get the names of the character vector `data_char_ukimmig2010`, and assign this to a document variable (*docvar*).
```{r}
docvars(my_corpus, "Party") <- names(data_char_ukimmig2010)
docvars(my_corpus, "Year") <- 2010
summary(my_corpus)
```

If we wanted to tag each document with additional meta-data not considered a document variable of interest for analysis, but rather something that we need to know as an attribute of the document, we could also 
add those to our corpus.
```{r}
metadoc(my_corpus, "language") <- "english"
metadoc(my_corpus, "docsource")  <- paste("data_char_ukimmig2010", 1:ndoc(my_corpus), sep = "_")
summary(my_corpus, showmeta = TRUE)
```

The last command, `metadoc`, allows you to define your own document meta-data fields.  Note that in assiging just the single value of `"english"`, R has recycled the value until it matches the number of documents in the corpus.  In creating
a simple tag for our custom metadoc field `docsource`, we used the quanteda function `ndoc()` to retrieve
the number of documents in our corpus.  This function is deliberately designed to work in a way similar to 
functions you may already use in R, such as `nrow()` and `ncol()`.

### Loading in files using the readtext package

```{r, eval=FALSE}
require(readtext)

# Twitter json
mytf1 <- readtext("~/Dropbox/QUANTESS/social media/zombies/tweets.json")
my_corpusTwitter <- corpus(mytf1)
summary(my_corpusTwitter, 5)
# generic json - needs a textfield specifier
mytf2 <- readtext("~/Dropbox/QUANTESS/Manuscripts/collocations/Corpora/sotu/sotu.json",
                  textfield = "text")
summary(corpus(mytf2), 5)
# text file
mytf3 <- readtext("~/Dropbox/QUANTESS/corpora/project_gutenberg/pg2701.txt", cache = FALSE)
summary(corpus(mytf3), 5)
# multiple text files
mytf4 <- readtext("~/Dropbox/QUANTESS/corpora/inaugural/*.txt", cache = FALSE)
summary(corpus(mytf4), 5)
# multiple text files with docvars from filenames
mytf5 <- readtext("~/Dropbox/QUANTESS/corpora/inaugural/*.txt", 
                  docvarsfrom = "filenames", sep = "-", docvarnames = c("Year", "President"))
summary(corpus(mytf5), 5)
# XML data
mytf6 <- readtext("~/Dropbox/QUANTESS/quanteda_working_files/xmlData/plant_catalog.xml", 
                  textfield = "COMMON")
summary(corpus(mytf6), 5)
# csv file
write.csv(data.frame(inaugSpeech = texts(data_corpus_inaugural), 
                     docvars(data_corpus_inaugural)),
          file = "/tmp/inaug_texts.csv", row.names = FALSE)
mytf7 <- readtext("/tmp/inaug_texts.csv", textfield = "inaugSpeech")
summary(corpus(mytf7), 5)
```


## How a quanteda corpus works

### Corpus principles

A corpus is designed to be a "library" of original documents that have been converted to plain, UTF-8 encoded text, and stored along with meta-data at the corpus level and at the document-level.  We have a special name for document-level meta-data: *docvars*.  These are variables or features that describe attributes of each document.

A corpus is designed to be a more or less static container of texts with respect to processing and analysis.  This means that the texts in corpus are not designed to be changed internally through (for example) cleaning or pre-processing steps, such as stemming or removing punctuation.  Rather, texts can be extracted from the corpus as part of processing, and assigned to new objects, but the idea is that the corpus will remain as an original reference copy so that other analyses -- for instance those in which stems and punctuation were required, such as analyzing a reading ease index -- can be performed on the same corpus.

To extract texts from a corpus, we use an extractor, called `texts()`.  

```{r}
texts(data_corpus_inaugural)[2]
```

To summarize the texts from a corpus, we can call a `summary()` method defined for a corpus.

```{r}
summary(data_corpus_irishbudget2010)
```

We can save the output from the summary command as a data frame, and plot some basic descriptive statistics with this information:

```{r, fig.width = 8}
tokenInfo <- summary(data_corpus_inaugural)
if (require(ggplot2))
    ggplot(data=tokenInfo, aes(x = Year, y = Tokens, group = 1)) + geom_line() + geom_point() +
        scale_x_continuous(labels = c(seq(1789, 2017, 12)), breaks = seq(1789, 2017, 12)) +
    theme_bw()

# Longest inaugural address: William Henry Harrison
tokenInfo[which.max(tokenInfo$Tokens), ] 
```


## Tools for handling corpus objects

### Adding two corpus objects together

The `+` operator provides a simple method for concatenating two corpus objects.  If they contain different sets of document-level variables, these will be stitched together in a fashion that guarantees that no information is lost.  Corpus-level medata data is also concatenated.

```{r}
library(quanteda)
my_corpus1 <- corpus(data_corpus_inaugural[1:5])
my_corpus2 <- corpus(data_corpus_inaugural[53:58])
my_corpus3 <- my_corpus1 + my_corpus2
summary(my_corpus3)
```

### Subsetting corpus objects

There is a method of the `corpus_subset()` function defined for corpus objects, where a new corpus can be extracted based on logical conditions applied to *docvars*:

```{r}
summary(corpus_subset(data_corpus_inaugural, Year > 1990))
summary(corpus_subset(data_corpus_inaugural, President == "Adams"))
```


## Exploring corpus texts

The `kwic` function (keywords-in-context) performs a search for a word and allows us to view the contexts in which it occurs:

```{r, tidy=TRUE}
kwic(data_corpus_inaugural, "terror")
```

```{r}
kwic(data_corpus_inaugural, "terror", valuetype = "regex")
```

```{r}
kwic(data_corpus_inaugural, "communist*")
```


In the above summary, `Year` and `President` are variables associated with each document. We can access such variables with the `docvars()` function.

```{r}
# inspect the document-level variables
head(docvars(data_corpus_inaugural))

# inspect the corpus-level metadata
metacorpus(data_corpus_inaugural)
```

More corpora are available from the [quanteda.corpora](http://github.com/quanteda/quanteda.corpora) package.

# Extracting Features from a Corpus

In order to perform statistical analysis such as document scaling, we must extract a matrix associating values for certain features with each document. In quanteda, we use the `dfm` function to produce such a matrix. "dfm" is short for *document-feature matrix*, and always refers to documents in rows and "features" as columns.  We fix this dimensional orientation because it is standard in data analysis to have a unit of analysis as a row, and features or variables
pertaining to each unit as columns.  We call them "features" rather than terms, because features are more general than terms: they can be defined as raw terms, stemmed terms, the parts of speech of terms, terms after stopwords have been removed, or a dictionary class to which a term belongs.  Features can be entirely general, such as ngrams or syntactic dependencies, and we leave this open-ended.

## Tokenizing texts

To simply tokenize a text, quanteda provides a powerful command called `tokens()`.  This produces an intermediate object, consisting of a list of tokens in the form of character vectors, where each element of the list corresponds to an input document.

`tokens()` is deliberately conservative, meaning that it does not remove anything from the text unless told to do so.

```{r}
txt <- c(text1 = "This is $10 in 999 different ways,\n up and down; left and right!", 
         text2 = "@kenbenoit working: on #quanteda 2day\t4ever, http://textasdata.com?page=123.")
tokens(txt)
tokens(txt, remove_numbers = TRUE,  remove_punct = TRUE)
tokens(txt, remove_numbers = FALSE, remove_punct = TRUE)
tokens(txt, remove_numbers = TRUE,  remove_punct = FALSE)
tokens(txt, remove_numbers = FALSE, remove_punct = FALSE)
tokens(txt, remove_numbers = FALSE, remove_punct = FALSE, remove_separators = FALSE)
```

We also have the option to tokenize characters:
```{r}
tokens("Great website: http://textasdata.com?page=123.", what = "character")
tokens("Great website: http://textasdata.com?page=123.", what = "character", 
         remove_separators = FALSE)
```

and sentences:
```{r}
# sentence level         
tokens(c("Kurt Vongeut said; only assholes use semi-colons.", 
           "Today is Thursday in Canberra:  It is yesterday in London.", 
           "En el caso de que no puedas ir con ellos, ¿quieres ir con nosotros?"), 
          what = "sentence")
```

## Constructing a document-feature matrix

Tokenizing texts is an intermediate option, and most users will want to skip straight to constructing
a document-feature matrix.  For this, we have a Swiss-army knife function, called `dfm()`, which performs
tokenization and tabulates the extracted features into a matrix of documents by features.  Unlike
the conservative approach taken by `tokens()`, the `dfm()` function applies certain options by default,
such as `tolower()` -- a separate function for lower-casing texts -- and removes punctuation.  All of the options to `tokens()` can be passed to `dfm()`, however.

```{r}
my_corpus <- corpus_subset(data_corpus_inaugural, Year > 1990)

# make a dfm
my_dfm <- dfm(my_corpus)
my_dfm[, 1:5]
```

Other options for a `dfm()` include removing stopwords, and stemming the tokens.
```{r}
# make a dfm, removing stopwords and applying stemming
myStemMat <- dfm(my_corpus, remove = stopwords("english"), stem = TRUE, remove_punct = TRUE)
myStemMat[, 1:5]
```

The option `remove` provides a list of tokens to be ignored.  Most users will 
supply a list of pre-defined "stop words", defined for numerous languages, accessed through 
the `stopwords()` function:
```{r}
head(stopwords("english"), 20)
head(stopwords("russian"), 10)
head(stopwords("arabic"), 10)
```

### Viewing the document-feature matrix

The dfm can be inspected in the Enviroment pane in RStudio, or by calling R's `View` function. Calling `plot` on a dfm will display a wordcloud.

```{r warning=FALSE, fig.width = 8, fig.height = 8}
my_dfm <- dfm(data_char_ukimmig2010, remove = stopwords("english"), remove_punct = TRUE)
my_dfm
```

To access a list of the most frequently occurring features, we can use `topfeatures()`:
```{r}
topfeatures(my_dfm, 20)  # 20 top words
```

Plotting a word cloud is done using `textplot_wordcloud()`, for a `dfm` class object.  This function passes arguments through to `wordcloud()` from the **wordcloud** package, and can prettify the plot using the same arguments:
```{r warning=FALSE, fig.width = 7, fig.height = 7}
set.seed(100)
textplot_wordcloud(my_dfm, min_count = 6, random_order = FALSE,
                   rotation = .25, 
                   color = RColorBrewer::brewer.pal(8,"Dark2"))
```

### Grouping documents by document variable 

Often, we are interested in analysing how texts differ according to substantive factors which may be encoded in the document variables, rather than simply by the boundaries of the document files. We can group documents which share the same value for a document variable when creating a dfm:

```{r}
by_party_dfm <- dfm(data_corpus_irishbudget2010, groups = "party", 
                  remove = stopwords("english"), remove_punct = TRUE)
```

We can sort this dfm, and inspect it:
```{r}
dfm_sort(by_party_dfm)[, 1:10]
```

### Grouping words by dictionary or equivalence class

For some applications we have prior knowledge of sets of words that are indicative of traits we would like to measure from the text. For example, a general list of positive words might indicate positive sentiment in a movie review, or we might have a dictionary of political terms which are associated with a particular ideological stance. In these cases, it is sometimes useful to treat these groups of words as equivalent for the purposes of analysis, and sum their counts into classes. 

For example, let's look at how words associated with terrorism and words associated with the economy vary by President in the inaugural speeches corpus. From the original corpus, we select Presidents since Clinton:

```{r}
recent_corpus <- corpus_subset(data_corpus_inaugural, Year > 1991)
```

Now we define a demonstration dictionary:
```{r}
my_dict <- dictionary(list(terror = c("terrorism", "terrorists", "threat"),
                          economy = c("jobs", "business", "grow", "work")))
```


We can use the dictionary when making the dfm:
```{r}
by_pres_mat <- dfm(recent_corpus, dictionary = my_dict)
by_pres_mat
```

The constructor function `dictionary()` also works with two common "foreign" dictionary formats: the 
LIWC and Provalis Research's Wordstat format.  For instance, we can load the LIWC and apply this to the Presidential inaugural speech corpus:
```{r, eval = FALSE}
liwcdict <- dictionary(file = "~/Dropbox/QUANTESS/dictionaries/LIWC/LIWC2001_English.dic",
                       format = "LIWC")
liwcdfm <- dfm(data_corpus_inaugural[52:58], dictionary = liwcdict)
liwcdfm[, 1:10]
```

# Further Examples

## Similarities between texts

```{r fig.width = 6}
pres_dfm <- dfm(corpus_subset(data_corpus_inaugural, Year > 1980), 
               remove = stopwords("english"), stem = TRUE, remove_punct = TRUE)
obama_simil <- textstat_simil(pres_dfm, c("2009-Obama" , "2013-Obama"), 
                             margin = "documents", method = "cosine")
obama_simil
# dotchart(as.list(obamaSimil)$"2009-Obama", xlab = "Cosine similarity")
```

We can use these distances to plot a dendrogram, clustering presidents:
```{r, fig.width = 10, fig.height = 7, eval = FALSE}
data(data_corpus_sotu, package = "quanteda.corpora")
pres_dfm <- dfm(corpus_subset(data_corpus_sotu, Date > as.Date("1980-01-01")), 
               stem = TRUE, remove_punct = TRUE,
               remove = stopwords("english"))
pres_dfm <- dfm_trim(pres_dfm, min_termfreq = 5, min_docfreq = 3)

# hierarchical clustering - get distances on normalized dfm
pres_dist_mat <- textstat_dist(dfm_weight(pres_dfm, "prop"))
# hiarchical clustering the distance object
pres_cluster <- hclust(pres_dist_mat)
# label with document names
pres_cluster$labels <- docnames(pres_dfm)
# plot as a dendrogram
plot(pres_cluster, xlab = "", sub = "", main = "Euclidean Distance on Normalized Token Frequency")
```

We can also look at term similarities:
```{r}
sim <- textstat_simil(pres_dfm, c("fair", "health", "terror"), method = "cosine", margin = "features")
lapply(as.list(sim), head, 10)
```

## Scaling document positions

Here is a demonstration of unsupervised document scaling comparing the "wordfish" model: 
```{r}
# make prettier document names
ie_dfm <- dfm(data_corpus_irishbudget2010)
textmodel_wordfish(ie_dfm, dir = c(2, 1))
```

## Topic models

**quanteda** makes it very easy to fit topic models as well, e.g.:

```{r}
quant_dfm <- dfm(data_corpus_irishbudget2010, 
                remove_punct = TRUE, remove_numbers = TRUE, remove = stopwords("english"))
quant_dfm <- dfm_trim(quant_dfm, min_termfreq = 4, max_docfreq = 10)
quant_dfm

set.seed(100)
if (require(topicmodels)) {
    my_lda_fit20 <- LDA(convert(quant_dfm, to = "topicmodels"), k = 20)
    get_terms(my_lda_fit20, 5)
}
```