---
title: "Quanteda Quickstart"
output:
  rmarkdown::html_document:
    theme: null
    css: mystyle.css
    toc: yes
vignette: >
  %\VignetteIndexEntry{Quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Quickstart}
-->

```{r echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```
## Installing quanteda

The code for the `quanteda` package currently resides on
<http://github/kbenoit/quanteda>.  From an Internet-connected
computer, you can install the package directly using the
`devtools` package:

```{r eval=FALSE}
library(devtools)
if (!require(quanteda)) install_github("kbenoit/quanteda")
```

This will download the package from github and install it on your computer.
For other branches, for instance if you wish to install the
development branch (containing work in progress) rather than the
master, you should instead run:

```{r eval=FALSE}
# to install the latest dev branch version quanteda from Github use:
install_github("kbenoit/quanteda", dependencies=TRUE, quick=TRUE, ref="dev")
```

Typically, the `dev` branch of a software package is under active
development --- so while it contains the latest updates, it is more likely
to have bugs. The `master` branch might be missing some of the newer
features, but should be more reliable.



## Working with a text corpus

To try the functions provided for interacting with corpora, load the `inaugCorpus` object packaged with `quanteda`. This corpus contains US presidents' inaugural addresses since 1789, with document-level variables for the year of each address (`Year`) and the last name of the president (`President`). The `summary` command gives a brief description of the corpus, and a summary of the first `n` documents: 

```{r}
# make sure quanteda is loaded and load the corpus of inaugural addresses
library(quanteda)
data(inaugCorpus)
summary(inaugCorpus, n=3)
````

In the above summary, `Year` and `President` are variables associated with each document. We can access such variables with the `docvars` command.

```{r}
# check the document-level variable names
names(docvars(inaugCorpus))

# list the first few values
head(docvars(inaugCorpus))

# check the corpus-level metadata
metacorpus(inaugCorpus)

```

Many more corpora are available in the [quantedaData](http://github.com/kbenoit/quantedaData) package.

## Document Feature Matrix

