---
title: "Quanteda Quickstart"
output:
  html_document:
    theme: null
    css: mystyle.css
    toc: yes
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Quickstart}
-->

```{r echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```
## Installing quanteda

The code for the `quanteda` package currently resides on
<http://github/kbenoit/quanteda>.  From an Internet-connected
computer, you can install the package directly using the
`devtools` package:

```{r eval=FALSE}
library(devtools)
if (!require(quanteda)) install_github("kbenoit/quanteda")
```

This will download the package from github and install it on your computer.
For other branches, for instance if you wish to install the
development branch (containing work in progress) rather than the
master, you should instead run:

```{r eval=FALSE}
# to install the latest dev branch version quanteda from Github use:
install_github("kbenoit/quanteda", dependencies=TRUE, quick=TRUE, ref="dev")
```

Typically, the `dev` branch of a software package is under active
development --- so while it contains the latest updates, it is more likely
to have bugs. The `master` branch might be missing some of the newer
features, but should be more reliable.