---
title: "Performance improvement in version 4.0"
author: Kohei Watanabe and Stefan Müller
output:  
  html_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    message = FALSE,
    comment = "##",
    fig.width = 8, 
    fig.height = 2, 
    dpi = 150
)
```

```{r, echo=FALSE, include=FALSE}
#data_corpus_guardian <- readRDS('/home/kohei/Dropbox/Public/data_corpus_guardian2016-10k.rds') 
data_corpus_guardian <- readRDS('C:/Users/watan/Dropbox/Public/data_corpus_guardian2016-10k.rds')
```

**quanteda** version 4.0 can process textual data significantly faster than its earlier versions thanks to the `tokens_xptr object` and a new glob pattern matching mechanism. More information on the features and advantages of the new xptr object are available in a [separate vignette](./articles/tokens.html).

We created the **quanteda3** package from version 3.2 and compared it with version 4.0 on a Windows laptop with AMD Ryzen 7 PRO processor (8 cores). We used sentences from 10,000 English-language news articles in this benchmarking.

We repeat the same operation using different versions of the same function to get the distribution of execution time. The result shows that the execution time of many v4.0 functions is about half of the version 3.2 functions.

```{r}
#devtools::install_github("quanteda/quanteda3")
require(quanteda)
require(ggplot2)

# create text corpus
corp <- corpus_reshape(data_corpus_guardian)

# tokenize corpus
toks <- tokens(corp, remove_punct = FALSE, remove_numbers = FALSE, 
               remove_symbols = FALSE)

# transform tokens object to tokens_xptr object
xtoks <- as.tokens_xptr(toks)

ndoc(toks) # the number of sentences
sum(ntoken(toks)) # the total number of tokens
```

## Tokenize a corpus

Although the v4.0 tokenizer is more flexible, its speed is roughly the same as v3.2 tokenizer. The shorter execution time of the version 4.0 is due to the faster removal of punctuation marks, numbers and symbols. We compare the performance by tokenizing the corpus with the `tokens()` function from **quanteda** version version 4.0 and 3.2 ("v4" and "v3" in the plots, respectively). 


```{r, echo=FALSE, include=FALSE}
# adjust custom ggplot2 theme to make it consistent with other vignettes
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r}
microbenchmark::microbenchmark(
    v3 = quanteda3::tokens(corp, remove_punct = TRUE, remove_numbers = TRUE, 
                           remove_symbols = TRUE),
    v4 = tokens(corp, remove_punct = TRUE, remove_numbers = TRUE, 
                remove_symbols = TRUE),
    times = 10
) %>% autoplot(log = FALSE)
```

## Modify tokens objects

`as.tokens_xptr()` is inserted before the v4 functions to keep the original `tokens_xptr` object intact.

```{r}
# generate n-grams
microbenchmark::microbenchmark(
    v3 = tokens_ngrams(toks),
    v4 = as.tokens_xptr(xtoks) %>% 
        tokens_ngrams(),
    times = 10
) %>% autoplot(log = FALSE)

# lookup dictionary keywords
microbenchmark::microbenchmark(
    v3 = quanteda3::tokens_lookup(toks, dictionary =  data_dictionary_LSD2015),
    v4 = as.tokens_xptr(xtoks) %>% 
        tokens_lookup(dictionary = data_dictionary_LSD2015),
    times = 10
) %>% autoplot(log = FALSE)


# remove stop words
microbenchmark::microbenchmark(
    v3 = quanteda3::tokens_remove(toks, pattern = stopwords("en"), padding = TRUE),
    v4 = as.tokens_xptr(xtoks) %>% 
        tokens_remove(pattern = stopwords("en"), padding = TRUE),
    times = 10
) %>% autoplot(log = FALSE)

# compound tokens
microbenchmark::microbenchmark(
    v3 = quanteda3::tokens_compound(toks,  pattern = "&", window = 1),
    v4 = as.tokens_xptr(xtoks) %>% 
        tokens_compound(pattern = "&", window = 1),
    times = 10
) %>% autoplot(log = FALSE)

# group sentences to articles
microbenchmark::microbenchmark(
    v3 = quanteda3::tokens_group(toks),
    v4 = tokens_group(xtoks),
    times = 10
) %>% autoplot(log = FALSE)
```

## Combine tokens objects

```{r}
# get first 5000 documents
toks1 <- head(toks, 5000)

# get last 5000 documents
toks2 <- tail(toks, 5000)

# transform both objects to tokens_xptr objects
xtoks1 <- as.tokens_xptr(toks1)
xtoks2 <- as.tokens_xptr(toks2)

# combine tokens objects
microbenchmark::microbenchmark(
    v3 = quanteda3:::c.tokens(toks1, toks2),
    v4 = quanteda:::c.tokens_xptr(xtoks1, xtoks2),
    times = 10
) %>% autoplot()
```

## Construct a document-feature matrix

We also compare the speed of constructing a document-feature matrix (DFM) using tokens objects.

```{r}
microbenchmark::microbenchmark(
    v3 = quanteda3::dfm(toks),
    v4 = dfm(xtoks),
    times = 10
) %>% autoplot()
```

## Simple pipeline: tokenize corpus and create document-feature matrix

```{r}
microbenchmark::microbenchmark(
    v3 = quanteda3::tokens(corp) %>% 
        quanteda3::tokens_remove(stopwords("en"), padding = TRUE) %>% 
        quanteda3::dfm(remove_padding = TRUE),
    v4 = tokens(corp, xptr = TRUE) %>% 
        tokens_remove(stopwords("en"), padding = TRUE) %>% 
        dfm(remove_padding = TRUE),
    times = 10
) %>% autoplot()
```

