---
title: "External pointer-based tokens objects"
author: Kohei Watanabe, Stefan Müller, and Kenneth Benoit
output:  
  html_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    message = FALSE,
    comment = "##",
    fig.width = 8, 
    fig.height = 2, 
    dpi = 150
)
```

## Introduction to `tokens_xptr` in **quanteda** 4

In **quanteda** version 4.0, we introduced a new class of tokens object, `tokens_xptr`. This object enhances the efficiency of adjusting tokens objects based on large text corpora.  A sub-class of the `tokens` object, `tokens_xptr` is built using on the external-pointer object `Rcpp::XPtr`.  External pointers allow **quanteda**'s tokens operations to pass references to the object to be modified by the highly optimized C++ routines that perform these operations, without the overhead of passing the objects by value back and forth from C++ to R code.

Prior to version 4, **quanteda**'s only version of a `tokens` object has been an R list of integer vectors with each integer mapped to a unique type.  With each call of a `tokens_*()` function, this object is converted to a C++ object, modified in C++, and then passed by value back to R and then recreated into an R `tokens` object. 
Such conversion between R and C++ consumes a large part of the execution time of each `tokens_*()` function call, and this time consumption grows as the size of the tokens object increases.

With the introduction of the `tokens_xptr` object, this conversion upon passing the object for modification is fundamentally changed, to passing just the reference to the tokens object so that it can be modified by passing a reference (the external pointer) to the object, eliminating the transaction costs of passing the classic `tokens` object by value.  The benefits of this are especially visible when large tokens objects are processed, say those with more than one million tokens, while still benefitting from the sequential steps used in the typical **quanteda** tokens formatting pipeline, such as creating the tokens, removing some based on a pattern, lower-casing them, forming n-grams, etc.  Formerly, with each step this large object was passed by value.  With `tokens_xptr`, successive token processing steps only pass the external pointer, resulting in significant efficiency gains without changing the clear syntax and workflow used in a typical **quanteda** tokens processing pipeline.

These efficiency gains are only evident for large tokens objects, for instance those with more than one million tokens.  The default `tokens` object format remains the non-external pointer type, but below, we show how to create and work with the `tokens_xptr` variant.  All of the `tokens_*()` functions work the same for both object classes, so that the user-facing differences are almost unnoticeable between the two object types.

A detailed comparison of performance gains and execution times for `tokens` versus `tokens_xptr` objects is available in a [separate article](./articles/pkgdown/benchmarks_xptr.html). 

```{r}
library("quanteda")
```


## Creating `tokens_xptr`` objects

There are two easy methods of creating a `tokens_xptr` object.

First, you can create a `tokens_xptr` from `tokens` using the `as.tokens_xptr()` method.
```{r}
toks <- tokens(data_corpus_inaugural)
print(toks, 0, 0)
xtoks <- as.tokens_xptr(toks)
print(xtoks, 0, 0)
```

Second, this step can be performed by setting `xptr = TRUE` argument in `tokens()`.
```{r}
xtoks2 <- tokens(data_corpus_inaugural, xptr = TRUE)
print(xtoks2, 0, 0)
```

When `xtoks` object is printed, the address of its underlying data is shown as `(pointer to xxxx)`. Because two different objects were created above, the addresses displayed for `xtoks` and `xtoks2` are different.  We expand on this below, in the discussion of shallow versus deep copies, when binding new object names to existing `toks_xptr` objects.


## Modifying `tokens_xptr` objects {#modifying-xptr}

### Shallow copies versus deep copies

When we create a new object by assigning that to an existing `tokens_xptr`, this creates what is known as a "shallow copy".  A _shallow copy_ is just a copy of the pointer to the memory address where the `tokens_xptr` version of the tokens objects exists.  This contrasts with a "deep copy", in which the new object takes its data through a physical copy of the data from the old memory location to a new one for the object.  Assignment via a _deep copy_ is the how assignment of a a regular `tokens` object --- and indeed most R objects --- works.

We can see this in operation here:
```{r}
toks <- data_corpus_inaugural |>
    corpus_reshape(to = "sentences") |>
    tokens()

xtoks <- as.tokens_xptr(toks)
print(xtoks, 0, 0)

# shallow copy
xtoks_copy <- xtoks
print(xtoks_copy, 0, 0)
```

As an external pointer object, the underlying data of the copy of the object `xtoks_copy` is the same. This means that operations on `xtoks` using `tokens_*` functions also affect `xtoks_copy` even after the copying (shallow copy). 

```{r}
print(xtoks, max_ndoc = 1)

xtoks_copy <- xtoks_copy |>
    tokens_remove(stopwords("en")) |>
    tokens_tolower()

print(xtoks, max_ndoc = 1)
```

This is not the normal R behaviour for modifying objects, but does exist in other packages that use external pointers, for instance, **data.table**, which has an entire [vignette on reference semantics](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-reference-semantics.html).

If you want to copy the underlying data of a `tokens_xptr` object (a _deep_ copy), this can be done explicitly using `copy()` or using `as.tokens_xptr()`.
```{r}
# deep copies
# print(copy(xtoks), 0, 0)
# print(as.tokens_xptr(xtoks), 0, 0)
```
Each of these has a different memory address from that of the `xtoks` object above.


### Modifying `tokens_xptr` objects using `tokens_*()` functions

The functions that modify `tokens_xptr` objects work in exactly the same way as they do for `tokens` objects, with the exception that they modify those objects by reference rather than passing copies by value back and forth from R to the underlying C++ code that performs the work.  

Not all of the `tokens_()` functions are implemented this way, because the efficiency gains for some would be negligible.  `tokens_select()`, `tokens_compound()`, `tokens_lookup()`, `tokens_ngrams()` and `tokens_replace()` will return a shallow copy of the input object for efficiency.  However, `tokens_subset()`, `[]`, `tokens_group()`, `tokens_segment()` and `tokens_chunk()` return a deep copy because they can change the number of the order of the documents.  If in any doubt, inspect the attention to the address of the pointer `(pointer to xxxx)` when using `tokens_xptr` objects.

```{r}
print(xtoks, 0, 0)

# shallow copy
xtoks2 <- tokens_remove(xtoks, stopwords("en"))
print(xtoks2, 0, 0)

# deep copy
xtoks3 <- tokens_subset(xtoks, lengths(toks) > 150)
print(xtoks3, 0, 0)
```


## Saving `tokens_xptr` objects

Because they point to a memory location rather than containing the data itself, a `tokens_xptr` object needs first needs to be converted to a `tokens` object before its data can be saved to disk.  This is simple and performed by applying the `as.tokens()` method to the `tokens_xptr` object.  After conversion, it can be saved normally using `save()` or `saveRDS()`.

```{r, eval=FALSE}
# converting to tokens before saving
save(as.tokens(xtoks), file = "~/tmp/xtoks.rda")
saveRDS(as.tokens(xtoks), file = "~/tmp/xtoks.rds")
```

## Creating a document-feature matrix from a `tokens_xptr` object

Passing a `tokens_xptr` object directly to `dfm()` to compile a document-feature matrix avoids a deep copy of that object and for larger objects can significantly reduce execution times for dfm creation. 

We demonstrate this below, using several`tokens_*()` functions to `tokens` and `tokens_xptr` objects to construct a `dfm` in this mock pipeline. The execution time for the latter is significantly shorter than for the former, thanks to the cost saved by passing the data without the expensive conversion.

```{r, echo=FALSE, include=FALSE}
# adjust custom ggplot2 theme to make it consistent with other vignettes
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r}
library("ggplot2")
library("microbenchmark")

data_corpus_guardian <- quanteda.corpora::download("data_corpus_guardian")
toks <- data_corpus_guardian |>
    corpus_reshape(to = "sentences") |>
    tokens()

microbenchmark::microbenchmark(
    tokens = toks |> 
        tokens_remove(stopwords("en"), padding = TRUE) |> 
        tokens_remove(c("http:*", "https:*", "*.com", "*.org", "*.net"), padding = TRUE) |> 
        tokens_compound(newsmap::data_dictionary_newsmap_en) |> 
        dfm(remove_padding = TRUE),
    tokens_xptr = as.tokens_xptr(toks) |> 
        tokens_remove(stopwords("en"), padding = TRUE) |> 
        tokens_remove(c("http:*", "https:*", "*.com", "*.org", "*.net"), padding = TRUE) |> 
        tokens_compound(newsmap::data_dictionary_newsmap_en) |> 
        dfm(remove_padding = TRUE),
    times = 10
) |> autoplot(log = FALSE)
```

We explore further benchmarks and comparisons for `tokens_xptr` usage in a [separate article](./articles/pkgdown/benchmarks_xptr.html).
