---
title: "Example: Latent Semantic Analysis (LSA)"
author: Haiyan Wang
output: 
  html_document:
    toc: true
    number_sections: true
---

In this vignette, we show how to perform Latent Semantic Analysis using the **quanteda** package based on Grossman and Frieder’s  [*Information Retrieval, Algorithms and Heuristics*](http://www1.se.cuhk.edu.hk/%7Eseem5680/lecture/LSI-Eg.pdf).

LSA decomposes document-feature matrix into a reduced vector space that is assumed to reflect semantic structure.

New documents or queries can be 'folded-in' to this constructed latent semantic space for downstream tasks. 

```{r, message = FALSE}
library(quanteda)
```

# Create a document-feature matrix

```{r}
txt <- c(d1="Shipment of gold damaged in a fire",
         d2="Delivery of silver arrived in a silver truck",
         d3="Shipment of gold arrived in a truck" )

mydfm <- dfm(txt)
mydfm
```

# Construct the LSA model
```{r}
mylsa <- textmodel_lsa(mydfm)
```

 the new document vector coordinates in the reduced 2-dimensional space is:
```{r}
mylsa$docs[, 1:2]
```

# Apply the constructed LSA model to new data
Now the new unseen document can be represented in the reduced 2-dimensional space.
The unseen query document:
```{r}
querydfm <- dfm(c("gold silver truck")) %>%
    dfm_select(pattern = mydfm)
querydfm
```
```{r}
newq <- predict(mylsa, querydfm)
newq$docs_newspace[, 1:2]
```
