---
title: "Example: Chinese text analysis"
author: "Haiyan Wang"
output: 
  html_document:
    toc: true
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "##"
)
```

```{r, message=FALSE, warnings=FALSE}
library("quanteda")
```

# Download corpus

Download corpus constructed from *Report on the Work of the Government* published by Premier of the State Council between 1954 and 2017. You can download the corpus using the **quanteda.corpora** package.

```{r, message=FALSE,eval=FALSE}
# read text files
remotes::install_github("quanteda/quanteda.corpora")
```

```{r}
library("quanteda.corpora")
corp <- quanteda.corpora::download(url = "https://www.dropbox.com/s/37ojd5knz1qeyul/data_corpus_chinesegovreport.rds?dl=1")
```

# Tokenization

```{r}
# Chinese stopwords
ch_stop <- stopwords("zh", source = "misc")

# tokenize
toks_ch <- corp |> 
    tokens(remove_punct = TRUE) |>
    tokens_remove(pattern = ch_stop)

# Construct a dfm
dfmat_ch <- dfm(toks_ch)

# Get most frequent features
topfeatures(dfmat_ch)
```

# Analysis

## Word cloud

```{r, fig.height = 5, fig.width = 5, fig.align = "center"}
# Plot a word cloud
set.seed(100)

# to set the font correctly for macOS
library("quanteda.textplots")
textplot_wordcloud(dfmat_ch, min_count = 500, random_order = FALSE,
                   rotation = .25, max_words = 100,
                   min_size = 0.5, max_size = 2.8,
                   font = if (Sys.info()['sysname'] == "Darwin") "SimHei" else NULL,
                   color = RColorBrewer::brewer.pal(8, "Dark2"))
```

## Feature co-occurrence matrix

```{r}
# fcm within the window size of 5
corp_ch17 <- corpus_subset(corp, Year == "2017")

toks_ch17 <- corp_ch17 |> 
    tokens(remove_punct = TRUE) |> 
    tokens_remove(ch_stop)
    
fcmat_ch <- corp_ch17 |> 
    tokens() |> 
    fcm(context = "window", window = 5)  

topfeatures(fcmat_ch["改革", ])
```

## Unsupervised document scaling

In this example, we run a Wordfish model to show how to apply an unspervised document scaling method to Chinese texts.

```{r fig2, fig.height = 4, fig.width = 6, fig.align = "center"}
library("quanteda.textmodels")

tmod_wf <- textmodel_wordfish(dfmat_ch)

y <- 1954:2017
y <- y[y <= 1964 | y >= 1975]
y <- y[!y %in% c(1963, 1961, 1962, 1976, 1977)]
plot(y, tmod_wf$theta, xlab = "Year", ylab = "Position")
```

## Collocations

```{r}
# bigrams cross the whole dataset
library("quanteda.textstats")

tstat_col_ch <- textstat_collocations(toks_ch, size = 2, min_count = 20)
knitr::kable(head(tstat_col_ch, 10))

# bigrams in 2017 report
tstat_col_ch17 <- textstat_collocations(toks_ch17, size = 2)
knitr::kable(head(tstat_col_ch17, 10))
```
