---
title: "Example: Latent Semantic Analysis (LSA)"
author: "Haiyan Wang"
output: 
  html_document:
    toc: true
    number_sections: true
---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "##")
```

In this vignette, we show how to perform Latent Semantic Analysis using the **quanteda** package based on Grossman and Frieder's  [*Information Retrieval, Algorithms and Heuristics*](http://www1.se.cuhk.edu.hk/%7Eseem5680/lecture/LSI-Eg.pdf). We can reduce the dimensions of document-feature matrix while keeping the essential semantic structure.

```{r, message=FALSE}
library("quanteda")
```

## Create a document-feature matrix

```{r}
txt <- c(d1 = "Shipment of gold damaged in a fire",
         d2 = "Delivery of silver arrived in a silver truck",
         d3 = "Shipment of gold arrived in a truck" )

dfmat <- txt |> 
    tokens() |> 
    dfm()
dfmat
```

## Fit the LSA model

```{r}
library("quanteda.textmodels")
tmod_lsa <- textmodel_lsa(dfmat)
```

The new document vector coordinates in the reduced 2-dimensional space is:

```{r}
tmod_lsa$docs[, 1:2]
```

## Apply the model to new data

Now the new unseen document can be represented in the reduced 2-dimensional space.

```{r}
dfmat_test <- tokens("gold silver truck") |> 
    dfm() |> 
    dfm_match(features = featnames(dfmat))
dfmat_test

pred_lsa <- predict(tmod_lsa, newdata = dfmat_test)
pred_lsa$docs_newspace[, 1:2]
```
