---
title: "Example: Convolutional Neural Network"
author: "Kohei Watanabe"
output: 
  html_document:
    toc: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "##", cache = TRUE)
```

In this vignette, we show how to implement neural networks using the **quanteda** and **torch** packages. In the version 4.5.0 of **quanteda**, we added multiple functions to make it an infrastructure for developing neural networks. This example focuses on a convolution neural network (CNN) but other types of models such as recurrent neural networks (RNN) and transformers can be also implemented with the aid of the **luz** package.

The code below is based on [an example](https://mlverse.github.io/luz/articles/examples/text-classification.html) published on **luz**'s website, but the CNN model is trained on **quanteda**'s tokens object. This makes the neural network models consistent with other text analyses tools and easier to integrate in existing pipelines.

```{r, message=FALSE}
library(quanteda)
library(torch)
library(luz)
```

## Download example data

Download the movie review texts with sentiment annotation, and decompress them in an temporary directory.

```{r}
tmp <- tempfile()
if (!file.exists(tmp))
    download.file("https://ai.stanford.edu/~amaas/data/sentiment/aclImdb_v1.tar.gz", tmp)
untar(tmp, exdir = tempdir(check = TRUE))
```

## Construct corpus

Extract information about the train/test split and sentiment labels from the file paths. HTML tags should be removed from the texts. 

```{r}
file <- list.files(file.path(tempdir(), "aclImdb"), recursive = TRUE, full.names = TRUE)
file <- stringi::stri_subset_regex(file, "(neg|pos)\\/\\d+_\\d+\\.txt")

txt <- sapply(file, function(f) paste0(readLines(f, warn = FALSE), collapse = "\n"))
txt <- stringi::stri_replace_all_fixed(txt, "<br />", " ")

m <- stringi::stri_match_first_regex(file, "(train|test)\\/(neg|pos)")
dat <- data.frame(text = txt, 
                  split = factor(m[,2], levels = c("train", "test")),
                  sentiment = match(m[,3], c("neg", "pos")) - 1)
corp <- corpus(dat)
```

## Tokenize texts

We tokenize texts using `tokens()`, remove only punctuations, and lower-case them using `tokens_tolower()`. To limit the complexity of the data, we apply `tokens_trim()` with the new `max_n` argument. This limits the size of the vocabulary (unique token types) to the top 20,000 most frequent words in the object.

```{r}
vocab_size <- 20000 # maximum size of the vocabulary

quanteda_options(verbose = TRUE)
toks <- tokens(corp, remove_punct = TRUE) %>% 
  tokens_tolower() %>% 
  tokens_trim(max_n = vocab_size)
```

## Define data loader

We use `dataset()` to feed data from the tokens object to neural network models. We can quickly access the documents in the tokens object using `as.matrix()` with the `extract` argument. The function returns the documents in the same length: when the documents are longer than `length`, the vectors are truncated; if they are shorter, the vectors are padded by zero. Since the indices in `torch_tensor` is one-based, we shift token IDs by adding one (the ID for padding is zero in tokens objects).

Importantly, before separating the tokens object into training and test sets using `tokens_subset()`, we have to convert it to a  [tokens_xptr](../tokens_xptr.html) object. This is not only for the quickly accessing the documents but also for preserving the original token IDs (`tokens_subset()` reassign tokens IDs when the list based tokens object is provided). Alternatively, we can match the token IDs between the training and test sets using `tokens_match()`.

```{r}
text_length <- 200 # maximum number length of the texts

movie_dataset <- dataset(
  name = "movie_dataset",
  
  initialize = function(data, text_length) {
    self$toks <- data
    self$dvars <- docvars(data)
  },
  .getitem = function(i) {
    list(
      x = as.matrix(self$toks, length = text_length, extract = i) + 1L,
      y = self$dvars$sentiment[i]
    )
  },
  .getbatch = function(i) {
    list(
      x = as.matrix(self$toks, length = text_length, extract = i, drop = FALSE) + 1L,
      y = self$dvars$sentiment[i]
      # alternatively
      #x = as.tensor(self$toks, length = text_length, extract = i),
      #y = torch_tensor(self$dvars$sentiment[i])
    )
  },
  .length = function() {
    ndoc(self$toks)
  }
)

xtoks <- as.tokens_xptr(toks) # important!
train_ds <- movie_dataset(tokens_subset(xtoks, split == "train"), text_length)
test_ds <- movie_dataset(tokens_subset(xtoks, split == "test"), text_length)
```

## Build CNN model

The code is identical to the original example except that `num_embeddings` is `vocab_size + 1L` for padding. The model starts with a module for word vectors in the embedding layer `nn_embedding`, followed by modules for convolution layers `nn_conv1d()`: the first layer captures the sequence of tokens such as phrases; the second layer further abstracts the occurrences sequences. Finally, the dense feed-forward network `nn_linear()` predicts the sentiment of documents. 

```{r}
embedding_dim <- 128 # size of the embedding vectors

model <- nn_module(
    initialize = function(vocab_size, embedding_dim) {
        self$embedding <- nn_sequential(
            nn_embedding(num_embeddings = vocab_size + 1L, embedding_dim = embedding_dim),
            nn_dropout(0.5)
        )
        
        self$convs <- nn_sequential(
            nn_conv1d(embedding_dim, 128, kernel_size = 7, stride = 3, padding = "valid"),
            nn_relu(),
            nn_conv1d(128, 128, kernel_size = 7, stride = 3, padding = "valid"),
            nn_relu(),
            nn_adaptive_max_pool2d(c(128, 1)) # reduces the length dimension
        )
        
        self$classifier <- nn_sequential(
            nn_flatten(),
            nn_linear(128, 128),
            nn_relu(),
            nn_dropout(0.5),
            nn_linear(128, 1)
        )
    },
    forward = function(x) {
        emb <- self$embedding(x)
        out <- emb$transpose(2, 3) %>% 
            self$convs() %>% 
            self$classifier()
        # we drop the last so we get (B) instead of (B, 1)
        out$squeeze(2)
    }
)

```

## Train the model

The hierarchical CNN model is trained over three iterations with hyper-parameters, `vocab_size` and `embedding_dim`. 

```{r}
fitted_model <- model %>% 
    setup(
        loss = nnf_binary_cross_entropy_with_logits,
        optimizer = optim_adam,
        metrics = luz_metric_binary_accuracy_with_logits()
    ) %>% 
    set_hparams(vocab_size = vocab_size, embedding_dim = embedding_dim) %>% 
    fit(train_ds, epochs = 3)
```

## Test the model

Evaluate the fitted model using the test set.

```{r}
fitted_model %>% 
    evaluate(test_ds) %>% 
    print()
```

