---
title: "Example: Convolutional Neural Network"
author: "Kohei Watanabe"
output: 
  html_document:
    toc: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

In this vignette, we show how to implement neural networks using the **quanteda** and **torch** packages. In the version 4.4.1 of **quanteda**, we added multiple functions to make it an infrastructure for developing neural networks. This example focuses on a convolution neural network (CNN) but other types of models such as recurrent neural networks (RNN) and transformers can be also implemented with the aid of the **luz** package.

The code below is based on [an example](https://mlverse.github.io/luz/articles/examples/text-classification.html) published in the website for **luz**, but the CNN model is trained on **quanteda**'s tokens object. This makes neural network models consistent with other analysis and even more efficient in textual data analysis projects.

```{r, message = FALSE}
library(quanteda)
library(torch)
library(luz)
```

## Download example data

Download the movie review texts with sentiment annotation, and decompress them in an temporary directory.

```{r}
tmp <- tempfile()
if (!file.exists(tmp))
    download.file("https://ai.stanford.edu/~amaas/data/sentiment/aclImdb_v1.tar.gz", tmp)
untar(tmp, exdir = tempdir(check = TRUE))
```

## Construct corpus

File paths contain information about train/test split and sentiment labels. HTML tags are also be removed from the texts. 

```{r}
file <- list.files(file.path(tempdir(), "aclImdb"), recursive = TRUE, full.names = TRUE)
file <- stringi::stri_subset_regex(file, "(neg|pos)\\/\\d+_\\d+\\.txt")

txt <- sapply(file, function(f) paste0(readLines(f, warn = FALSE), collapse = "\n"))
txt <- stringi::stri_replace_all_fixed(txt, "<br />", " ")

m <- stringi::stri_match_first_regex(file, "(train|test)\\/(neg|pos)")
dat <- data.frame(text = txt, 
                  split = factor(m[,2], levels = c("train", "test")),
                  sentiment = match(m[,3], c("neg", "pos")) - 1)
corp <- corpus(dat)
```

## Tokenize texts

We tokenize texts using `tokens()` and lower-case them using `tokens_tolower()`. To limit the complexity of the data, we apply `tokens_trim()` with the new `max_n` argument. This limits the size of the vocabulary (unique token types) to the top 20,000 most frequent words in the object.

```{r}
vocab_size <- 20000 # maximum size of the vocabulary

toks <- tokens(corp, remove_punct = TRUE, verbose = TRUE) %>% 
  tokens_tolower() %>% 
  tokens_trim(max_n = vocab_size)
```

## Define data loader

We use `dataset()` to feed data from the tokens object to neural network models. We can directly access to the documents in the tokens object using `as.matrix()` with the `extract` argument. The function returns all the documents in the same length: when the documents are longer than `length`, the vectors are truncated; if they are shorter, the vectors are padded by zero. Since the indices in `torch_tensor` is one-based, we shift token IDs by adding one (the ID for padding is zero in tokens objects).

Importantly, before separating the tokens object into training and test sets using `tokens_subset()`, we have to convert it to a  [tokens_xptr](../tokens_xptr.html) object. This in not only for accessing the documents more quickly but also for preserving the original token IDs. Alternatively, we can match the token IDs between the training and test sets using `tokens_match()`.

```{r}
text_length <- 200 # maximum number length of the texts

movie_dataset <- dataset(
  name = "movie_dataset",
  
  initialize = function(data, text_length) {
    self$toks <- data
    self$dvars <- docvars(data)
  },
  .getitem = function(i) {
    list(
      x = as.matrix(self$toks, length = text_length, extract = i) + 1L,
      y = self$dvars$sentiment[i]
    )
  },
  .getbatch = function(i) {
    list(
      x = as.matrix(self$toks, length = text_length, extract = i, drop = FALSE) + 1L,
      y = self$dvars$sentiment[i]
      # alternatively
      #x = as.tensor(self$toks, length = text_length, extract = i),
      #y = torch_tensor(self$dvars$sentiment[i])
    )
  },
  .length = function() {
    ndoc(self$toks)
  }
)

xtoks <- as.tokens_xptr(toks)
train_ds <- movie_dataset(tokens_subset(xtoks, split == "train"), text_length)
test_ds <- movie_dataset(tokens_subset(xtoks, split == "test"), text_length)
```

## Build CNN model

The code is identical to the original except that `num_embeddings` is `vocab_size + 1L` for padding. The model starts with a module for word vectors in the embedding layer `nn_embedding`, followed by modules convolution layers `nn_conv1d()`: the first layer captures the sequence tokens such as phrases; the second layer further abstract the occurrences of groups of sequences. Finally, the dense feed-forward network `nn_linear()` predict the sentiment of documents. 

```{r}
embedding_dim <- 128 # size of the embedding vectors

model <- nn_module(
    initialize = function(vocab_size, embedding_dim) {
        self$embedding <- nn_sequential(
            nn_embedding(num_embeddings = vocab_size + 1L, embedding_dim = embedding_dim),
            nn_dropout(0.5)
        )
        
        self$convs <- nn_sequential(
            nn_conv1d(embedding_dim, 128, kernel_size = 7, stride = 3, padding = "valid"),
            nn_relu(),
            nn_conv1d(128, 128, kernel_size = 7, stride = 3, padding = "valid"),
            nn_relu(),
            nn_adaptive_max_pool2d(c(128, 1)) # reduces the length dimension
        )
        
        self$classifier <- nn_sequential(
            nn_flatten(),
            nn_linear(128, 128),
            nn_relu(),
            nn_dropout(0.5),
            nn_linear(128, 1)
        )
    },
    forward = function(x) {
        emb <- self$embedding(x)
        out <- emb$transpose(2, 3) %>% 
            self$convs() %>% 
            self$classifier()
        # we drop the last so we get (B) instead of (B, 1)
        out$squeeze(2)
    }
)

```

## Train the model

The hierarchical CNN model is trained trained over three iterations with hyper-parameters, `vocab_size` and `embedding_dim`. 

```{r}
fitted_model <- model %>% 
    setup(
        loss = nnf_binary_cross_entropy_with_logits,
        optimizer = optim_adam,
        metrics = luz_metric_binary_accuracy_with_logits()
    ) %>% 
    set_hparams(vocab_size = vocab_size, embedding_dim = embedding_dim) %>% 
    fit(train_ds, epochs = 3)
```

## Test the model

Evaluate the fitted model using the test set.

```{r}
fitted_model %>% 
    evaluate(test_ds) %>% 
    print()
```

