---
title: "Quick Start Guide"
output: 
  html_document:
    toc: true
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = FALSE, 
                      comment = "##")
```

このビネットでは、**quanteda**の特徴と機能の基本的な概要を解説します。 追加のビネットについては、[ quanteda.ioにあるその他の記事](http://docs.quanteda.io/articles/index.html)を参照してください。

# パッケージのインストール

**quanteda**は[CRAN](https://CRAN.R-project.org/package=quanteda)からインストールできます。GUIのRパッケージインストーラを使用してインストールするか、次のコマンドを実行します。

```{r, eval = FALSE}
install.packages("quanteda") 
```

GitHubから最新の開発バージョンをインストールする方法については、https://github.com/kbenoit/quanteda を参照してください。

## インストールが推奨されるパッケージ

**quanteda**には**quanteda**と連携して機能を拡張する、一連のパッケージがあります。それらもインストールすることをお勧めします：

* [**readtext**](https://github.com/kbenoit/readtext)：多くの入力形式からテキストデータをRに簡単に読み込むパッケージ。
* [**spacyr**](https://github.com/kbenoit/spacyr)：pythonの[spaCy](http://spacy.io)ライブラリを使用した自然言語解析のためのパッケージ。品詞タグ付け、固有表現抽出、および係り受け関係の解析など。
* [**quantedaData**](https://github.com/kbenoit/quantedaData)：**quanteda**のビネット等で使用する追加のテキストデータ。

    ```{r eval = FALSE}
    devtools::install_github("kbenoit/quantedaData")
    ```
*  [**LIWCalike**](https://github.com/kbenoit/LIWCalike): [Linguistic Inquiry and Word Count](http://liwc.wpengine.com) (LIWC) アプローチによるテキスト分析のR実装.
    ```{r eval = FALSE}
    devtools::install_github("kbenoit/LIWCalike")
    ```

# コーパスの作成

まず、パッケージをロードして、パッケージ内の関数とデータにアクセスできるようにします。

```{r, message = FALSE}
library(quanteda)
```

## 利用可能なコーパス

**quanteda**にはテキストを読み込むためのシンプルで強力なコンパニオンパッケージ、[**readtext**](https://github.com/kbenoit/readtext)があります。
このパッケージのmain関数`readtext()`は、ディスクやURLからファイルやファイルセットを取り出し、**quanteda**の`corpus()`コンストラクタ関数から直接呼び出せる形式の、data.frameを返します。

`readtext()`で利用可能なファイルやデータの形式:

* テキスト(`.txt`)ファイル
* コンマ区切り値(`.csv`)ファイル
* XML形式のデータ
* JSON形式のFacebook APIのデータ
* JSON形式のTwitter APIのデータ
* より一般的なJSONデータ

**quanteda**のコーパスコンストラクタコマンド `corpus()`は、次のクラスのオブジェクトを受け付けます：

* キャラクタオブジェクトのベクトル(例:他のツールを使用してすでにワークスペースにロードしたテキスト)
* **tm**パッケージの `VCorpus`コーパスオブジェクト
* テキスト列と他のドキュメントレベルのメタデータを含むdata.frame

### 例：文字列からコーパスを作成

最も単純なケースは、すでにRのメモリにあるテキストのベクトルからコーパスを作成することです。テキストのベクトルをRに取り込む方法は実にたくさんあるので、高度なRユーザーは、入力するテキストをいろいろな方法で作り出せます。

すでにテキストベクトル形式のテキストがある場合は、コーパスコンストラクタ関数を直接呼び出すことができます。 次の例では、**quanteda**パッケージと一緒に配布される、イギリスの政党の2010年の選挙マニフェストから抽出されたテキストオブジェクト(`data_char_ukimmig2010`)をつかって、コーパスを作成しています。

```{r}
myCorpus <- corpus(data_char_ukimmig2010)  # テキストからコーパスを作成
summary(myCorpus)
```

コーパスを作成したあとで、もし必要ならば、ドキュメントレベルの変数(`docvars`と呼ばれるもの)をこのコーパスに追加することができます。

たとえば、Rの`names()`関数を使って文字ベクトル`data_char_ukimmig2010`の名前を取得し、これをドキュメント変数(`docvar`)に追加することができます。

```{r}
docvars(myCorpus, "Party") <- names(data_char_ukimmig2010)
docvars(myCorpus, "Year") <- 2010
summary(myCorpus)
```

分析の対象となるドキュメント変数では必ずしもないけれども、ドキュメントの属性として残しておきたいと思う追加のメタデータも、`docvars()`を使って、コーパスに追加することができます。

```{r}
metadoc(myCorpus, "language") <- "english"
metadoc(myCorpus, "docsource")  <- paste("data_char_ukimmig2010", 1:ndoc(myCorpus), sep = "_")
summary(myCorpus, showmeta = TRUE)
```

このコマンド `metadoc`では、ドキュメントメタデータフィールドを自由に定義することができます。 `language`フィールドに、単一の値`english`をアサインするときには、Rはコーパス内の文書の数と一致するまで値を再利用していることに注意してください。
独自のmetadocフィールド`docsource`のタグを作成する際には、quanteda関数`ndoc()`を使ってコーパスの文書数を取り出しました。この関数`ndoc()`は、`nrow()`や `ncol()`のようなR標準の関数と同じような方法で動作するように意図して設計されています。

### 例：**readtext**パッケージを用いたファイルの読み込み

```{r, eval=FALSE}
require(readtext)

# Twitter json
mytf1 <- readtext("~/Dropbox/QUANTESS/social media/zombies/tweets.json")
myCorpusTwitter <- corpus(mytf1)
summary(myCorpusTwitter, 5)
# generic json - needs a textfield specifier
mytf2 <- readtext("~/Dropbox/QUANTESS/Manuscripts/collocations/Corpora/sotu/sotu.json",
                  textfield = "text")
summary(corpus(mytf2), 5)
# text file
mytf3 <- readtext("~/Dropbox/QUANTESS/corpora/project_gutenberg/pg2701.txt", cache = FALSE)
summary(corpus(mytf3), 5)
# multiple text files
mytf4 <- readtext("~/Dropbox/QUANTESS/corpora/inaugural/*.txt", cache = FALSE)
summary(corpus(mytf4), 5)
# multiple text files with docvars from filenames
mytf5 <- readtext("~/Dropbox/QUANTESS/corpora/inaugural/*.txt", 
                  docvarsfrom = "filenames", sep = "-", docvarnames = c("Year", "President"))
summary(corpus(mytf5), 5)
# XML data
mytf6 <- readtext("~/Dropbox/QUANTESS/quanteda_working_files/xmlData/plant_catalog.xml", 
                  textfield = "COMMON")
summary(corpus(mytf6), 5)
# csv file
write.csv(data.frame(inaugSpeech = texts(data_corpus_inaugural), 
                     docvars(data_corpus_inaugural)),
          file = "/tmp/inaug_texts.csv", row.names = FALSE)
mytf7 <- readtext("/tmp/inaug_texts.csv", textfield = "inaugSpeech")
summary(corpus(mytf7), 5)
```

## コーパスオブジェクトの使い方

### コーパスの原理

コーパスは、元になったドキュメントを、プレーンなUTF-8でエンコードされたテキストに変換し、コーパスレベルおよびドキュメントレベルのメタデータと一緒に格納する、ドキュメントの「ライブラリ」として設計されています。ドキュメントレベルのメタデータには*docvars*という名前がつけられています。*docvars*は、各文書の属性を記述する変数（あるいは特徴）です。

コーパスは基本的には、処理および分析に関して、テキストの静的なコンテナになるように設計されています。これが意味するのは、コーパス内のテキストは、ステミングや句読点の削除などのテキストのクリーニングや前処理ステップによって内部的に変更されるようには設計されていない、ということです。このことで、処理の一部としてコーパスからテキストを抽出して新しいオブジェクトに割り当てた後でも、コーパスは元のテキストの参照コピーとして残り、たとえばステミング前の単語や句読点を必要とする可読性の分析などの分析を、同じコーパス上で実行することができるようになるのです。

コーパスからテキストを抽出するためには、`texts()`と呼ばれる抽出子を使用します。

```{r}
texts(data_corpus_inaugural)[2]
```


`summary()`メソッドにより、コーパス内のテキストの要約を行うことができます。

```{r}
summary(data_corpus_irishbudget2010)
```

summaryコマンドの出力をデータフレームとして保存し、この情報を使って基本的な記述統計をプロットすることができます：

```{r, fig.width = 8}
tokenInfo <- summary(data_corpus_inaugural)
if (require(ggplot2))
    ggplot(data=tokenInfo, aes(x = Year, y = Tokens, group = 1)) + geom_line() + geom_point() +
        scale_x_discrete(labels = c(seq(1789,2012,12)), breaks = seq(1789,2012,12) ) 

# Longest inaugural address: William Henry Harrison
tokenInfo[which.max(tokenInfo$Tokens), ] 
```


## コーパスに対する操作

### コーパスの結合

`+`演算子で、2つのコーパスオブジェクトを簡単に連結できます。ドキュメントレベルの変数が2つのコーパスで異なる場合には、ドキュメント変数は継ぎ接ぎされて、情報が失われることを防ぎます。コーパスレベルのメタデータも結合されます。

```{r}
library(quanteda)
mycorpus1 <- corpus(data_corpus_inaugural[1:5])
mycorpus2 <- corpus(data_corpus_inaugural[53:58])
mycorpus3 <- mycorpus1 + mycorpus2
summary(mycorpus3)
```

### コーパス内の文書の一部の文書の抽出

コーパスオブジェクトに対して定義された `corpus_subset()`関数のメソッドがあります。この関数により、ドキュメント変数
docvarsに適用される論理条件に基づいて文書を抽出します。

```{r}
summary(corpus_subset(data_corpus_inaugural, Year > 1990))
summary(corpus_subset(data_corpus_inaugural, President == "Adams"))
```


## コーパス内の文書の探索

`kwic`(KeyWords-In-Context)関数は単語の検索を行い、その単語が生起する文脈を表示します：

```{r, tidy=TRUE}
kwic(data_corpus_inaugural, "terror")
```

```{r}
kwic(data_corpus_inaugural, "terror", valuetype = "regex")
```

```{r}
kwic(data_corpus_inaugural, "communist*")
```


In the above summary, `Year` and `President` are variables associated with each document. We can access such variables with the `docvars()` function.

上記の要約では、「Year」と「President」は各文書に結び付けられた変数です。 `docvars()`関数でこのような変数にアクセスできます。

```{r}
# inspect the document-level variables
head(docvars(data_corpus_inaugural))

# inspect the corpus-level metadata
metacorpus(data_corpus_inaugural)
```

[quantedaData](http://github.com/kbenoit/quantedaData)パッケージをインストールすることにより、もっとたくさんのコーパスを試すことができます。




# Extracting Features from a Corpus

In order to perform statistical analysis such as document scaling, we
must extract a matrix associating values for certain features with each
document. In quanteda, we use the `dfm` function to produce such a matrix.  "dfm" is short for *document-feature matrix*, and always refers to documents
in rows and "features" as columns.  We fix this dimensional orientation because it is 
standard in data analysis to have a unit of analysis as a row, and features or variables
pertaining to each unit as columns.  We call them "features" rather than terms, because
features are more general than terms: they can be defined as raw terms, stemmed terms, the parts of speech of terms, terms after stopwords have been removed,
or a dictionary class to which a term belongs.  Features can be entirely general, such as ngrams or syntactic dependencies, and we leave this open-ended.

# コーパスから特徴を抽出

文書スケーリングなどの統計分析を行うために、特定のフィーチャの値をそれぞれに関連付ける行列を抽出する必要があります資料。 quantedaでは、このような行列を生成するために `dfm`関数を使います。 "dfm"は*document-feature matrix *の略で、常にドキュメントを参照します
列には「フィーチャ」が表示されます。 この次元の向きは固定されています
データ分析の標準であり、分析単位を行として持ち、特徴または変数
列として各ユニットに関連する。 我々はそれらを用語ではなく「機能」と呼んでいる。
機能は用語よりも一般的です：それらは、生の用語、ステムされた用語、用語の品詞、ストップワードが削除された後の用語、
または用語が属する辞書クラスである。 機能は、ngramsや構文依存性などの一般的なものにすることができ、私たちはこれをオープンエンドにします。

## Tokenizing texts

To simply tokenize a text, quanteda provides a powerful command called `tokens()`.  This produces an 
intermediate object, consisting of a list of tokens in the form of character vectors, where each element
of the list corresponds to an input document.

`tokens()` is deliberately conservative, meaning that it does not remove anything from the text unless
told to do so.

## 文書のトークン化

単純にテキストをトークン化するために、quantedaは `tokens()`と呼ばれる強力なコマンドを提供します。 これにより、中間オブジェクト。文字ベクトルの形式のトークンのリストで構成され、各要素リストのうちの1つは入力文書に対応する。

`tokens()`は故意に保守的です。つまり、そうするように言われた。

```{r}
txt <- c(text1 = "This is $10 in 999 different ways,\n up and down; left and right!", 
         text2 = "@kenbenoit working: on #quanteda 2day\t4ever, http://textasdata.com?page=123.")
tokens(txt)
tokens(txt, remove_numbers = TRUE,  remove_punct = TRUE)
tokens(txt, remove_numbers = FALSE, remove_punct = TRUE)
tokens(txt, remove_numbers = TRUE,  remove_punct = FALSE)
tokens(txt, remove_numbers = FALSE, remove_punct = FALSE)
tokens(txt, remove_numbers = FALSE, remove_punct = FALSE, remove_separators = FALSE)
```

We also have the option to tokenize characters:
```{r}
tokens("Great website: http://textasdata.com?page=123.", what = "character")
tokens("Great website: http://textasdata.com?page=123.", what = "character", 
         remove_separators = FALSE)
```

and sentences:
```{r}
# sentence level         
tokens(c("Kurt Vongeut said; only assholes use semi-colons.", 
           "Today is Thursday in Canberra:  It is yesterday in London.", 
           "En el caso de que no puedas ir con ellos, ¿quieres ir con nosotros?"), 
          what = "sentence")
```

## Constructing a document-feature matrix

Tokenizing texts is an intermediate option, and most users will want to skip straight to constructing
a document-feature matrix.  For this, we have a Swiss-army knife function, called `dfm()`, which performs
tokenization and tabulates the extracted features into a matrix of documents by features.  Unlike
the conservative approach taken by `tokens()`, the `dfm()` function applies certain options by default,
such as `toLower()` -- a separate function for lower-casing texts -- and removes punctuation.  All of the options to `tokens()` can be passed to `dfm()`, however.

## 文書行列の作成

テキストをトークン化することは中間的な選択肢であり、ほとんどのユーザーはスキップして直感的に
ドキュメントフィーチャマトリックス このために、私たちは `dfm()`と呼ばれるスイス軍ナイフ機能を持っています。
抽出された特徴をフィーチャによって文書のマトリックスに集計する。 とは異なり
`tokens()`、 `dfm()`関数が保守的に取り組んでいるアプローチは、デフォルトでいくつかのオプションを適用しています。
`toLower()` - 小文字のテキストのための別の関数 - や句読点を削除するなど)。 ただし、 `tokens()`のオプションはすべて `dfm()`に渡すことができます。

```{r}
myCorpus <- corpus_subset(data_corpus_inaugural, Year > 1990)

# make a dfm
myDfm <- dfm(myCorpus)
myDfm[, 1:5]
```

Other options for a `dfm()` include removing stopwords, and stemming the tokens.

`dfm()`のその他のオプションには、ストップワードの削除やトークンのステミングが含まれます。

```{r}
# make a dfm, removing stopwords and applying stemming
myStemMat <- dfm(myCorpus, remove = stopwords("english"), stem = TRUE, remove_punct = TRUE)
myStemMat[, 1:5]
```

The option `remove` provides a list of tokens to be ignored.  Most users will 
supply a list of pre-defined "stop words", defined for numerous languages, accessed through 
the `stopwords()` function:

`remove`オプションは、無視されるトークンのリストを提供します。 ほとんどのユーザーは
多数の言語に定義された定義済みの「ストップワード」のリストを、
`stopwords()`関数：

```{r}
head(stopwords("english"), 20)
head(stopwords("russian"), 10)
head(stopwords("arabic"), 10)
```



### Viewing the document-feature matrix

The dfm can be inspected in the Enviroment pane in RStudio, or by calling R's `View` function. Calling `plot` on a dfm will display a wordcloud using the [wordcloud package](https://cran.r-project.org/web/packages/wordcloud)

### 文書行列の表示

dfmは、RStudioの環境ペインで、またはRの `View`関数を呼び出すことで調べることができます。 dfmで `plot`を呼び出すと、[wordcloudパッケージ](https://cran.r-project.org/web/packages/wordcloud)を使ってワードクラウドが表示されます。

```{r warning=FALSE, fig.width = 8, fig.height = 8}
mydfm <- dfm(data_char_ukimmig2010, remove = stopwords("english"), remove_punct = TRUE)
mydfm
```

To access a list of the most frequently occurring features, we can use `topfeatures()`:
```{r}
topfeatures(mydfm, 20)  # 20 top words
```

Plotting a word cloud is done using `textplot_wordcloud()`, for a `dfm` class object.  This function passes arguments through to `wordcloud()` from the **wordcloud** package, and can prettify the plot using the same arguments:
```{r warning=FALSE, fig.width = 7, fig.height = 7}
set.seed(100)
textplot_wordcloud(mydfm, min.freq = 6, random.order = FALSE,
                   rot.per = .25, 
                   colors = RColorBrewer::brewer.pal(8,"Dark2"))
```

### Grouping documents by document variable 

Often, we are interested in analysing how texts differ according to substantive factors which may be encoded in the document variables, rather than simply by the boundaries of the document files. We can group documents which share the same value for a document variable when creating a dfm:

### 文書変数による文書のグループ化

多くの場合、文書ファイルの境界線ではなく、文書変数でエンコードされる可能性のある実質的要因によってテキストがどのように異なるかを分析することに興味があります。 dfmを作成するときに、ドキュメント変数と同じ値を共有するドキュメントをグループ化できます。

```{r}
byPartyDfm <- dfm(data_corpus_irishbudget2010, groups = "party", 
                  remove = stopwords("english"), remove_punct = TRUE)
```

We can sort this dfm, and inspect it:
```{r}
dfm_sort(byPartyDfm)[, 1:10]
```

Note that the most frequently occurring feature is "will", a word usually on English stop lists, but
one that is not included in quanteda's built-in English stopword list.  

最も頻繁に出現する特徴は、「英語」のストップリストにある単語である「意志」ですが、
quantedaの組み込み英語ストップワードリストには含まれていません。

### Grouping words by dictionary or equivalence class

For some applications we have prior knowledge of sets of words that are indicative of traits we would like to measure from the text. For example, a general list of positive words might indicate positive sentiment in a movie review, or we might have a dictionary of political terms which are associated with a particular ideological stance. In these cases, it is sometimes useful to treat these groups of words as equivalent for the purposes of analysis, and sum their counts into classes. 

For example, let's look at how words associated with terrorism and words associated with the economy vary by President in the inaugural speeches corpus. From the original corpus, we select Presidents since Clinton:

### 辞書を用いた語のグループ化

いくつかのアプリケーションでは、テキストから測定したい形質の指標である単語のセットについての事前知識があります。 たとえば、肯定的な言葉の一般的なリストは、映画レビューの肯定的な感情を示すかもしれない、または我々は特定のイデオロギー的なスタンスに関連付けられている政治用語の辞書を持っている可能性があります。 このような場合、分析目的でこれらの単語のグループを同等のものとして扱い、カウントをクラスに合計すると便利な場合があります。

たとえば、テロリズムに関連する言葉や経済に関連する言葉が、演説コーパスで大統領によってどのように異なるかを見てみましょう。 元のコーパスからは、クリントン以来の大統領選を選びます：

```{r}
recentCorpus <- corpus_subset(data_corpus_inaugural, Year > 1991)
```

Now we define a demonstration dictionary:
```{r}
myDict <- dictionary(list(terror = c("terrorism", "terrorists", "threat"),
                          economy = c("jobs", "business", "grow", "work")))
```


We can use the dictionary when making the dfm:
```{r}
byPresMat <- dfm(recentCorpus, dictionary = myDict)
byPresMat
```

The constructor function `dictionary()` also works with two common "foreign" dictionary formats: the 
LIWC and Provalis Research's Wordstat format.  For instance, we can load the LIWC and apply this to the Presidential inaugural speech corpus:

コンストラクタ関数 `dictionary()`は、2つの一般的な "外部"辞書形式でも動作します。
LIWCとProvalis ResearchのWordstat形式。 たとえば、LIWCを読み込み、これを大統領就任演説コーパスに適用することができます。

```{r, eval = FALSE}
liwcdict <- dictionary(file = "~/Dropbox/QUANTESS/dictionaries/LIWC/LIWC2001_English.dic",
                       format = "LIWC")
liwcdfm <- dfm(data_corpus_inaugural[52:58], dictionary = liwcdict)
liwcdfm[, 1:10]
```



# Further Examples

## Similarities between texts

# 追加の事例

## 文書の類似性

```{r fig.width = 6}
presDfm <- dfm(corpus_subset(data_corpus_inaugural, Year > 1980), 
               remove = stopwords("english"), stem = TRUE, remove_punct = TRUE)
obamaSimil <- textstat_simil(presDfm, c("2009-Obama" , "2013-Obama"), 
                             margin = "documents", method = "cosine")
obamaSimil
# dotchart(as.list(obamaSimil)$"2009-Obama", xlab = "Cosine similarity")
```


We can use these distances to plot a dendrogram, clustering presidents:
```{r, fig.width = 10, fig.height = 7, eval = FALSE}
data(data_corpus_SOTU, package = "quantedaData")
presDfm <- dfm(corpus_subset(data_corpus_SOTU, Date > as.Date("1980-01-01")), 
               stem = TRUE, remove_punct = TRUE,
               remove = stopwords("english"))
presDfm <- dfm_trim(presDfm, min_count = 5, min_docfreq = 3)
# hierarchical clustering - get distances on normalized dfm
presDistMat <- textstat_dist(dfm_weight(presDfm, "relfreq"))
# hiarchical clustering the distance object
presCluster <- hclust(presDistMat)
# label with document names
presCluster$labels <- docnames(presDfm)
# plot as a dendrogram
plot(presCluster, xlab = "", sub = "", main = "Euclidean Distance on Normalized Token Frequency")
```
(try it!)

We can also look at term similarities:
```{r}
sim <- textstat_simil(presDfm, c("fair", "health", "terror"), method = "cosine", margin = "features")
lapply(as.list(sim), head, 10)
```

## Scaling document positions

We have a lot of development work to do on the `textmodel()` function, but here is a demonstration of unsupervised document scaling comparing the "wordfish" model: 

## 文書の位置の特定

`textmodel()`関数にはたくさんの開発作業がありますが、ここでは "wordfish"モデルを比較した教師なしの文書スケーリングのデモンストレーションがあります：

```{r}
# make prettier document names
ieDfm <- dfm(data_corpus_irishbudget2010)
textmodel_wordfish(ieDfm, dir = c(2, 1))
```

## Topic models

**quanteda** makes it very easy to fit topic models as well, e.g.:

## トピックモデル

**quanteda**を使用すると、トピックモデルにも簡単にフィットできます(例：

```{r}
quantdfm <- dfm(data_corpus_irishbudget2010, 
                remove_punct = TRUE, remove_numbers = TRUE, remove = stopwords("english"))
quantdfm <- dfm_trim(quantdfm, min_count = 4, max_docfreq = 10)
quantdfm

if (require(topicmodels)) {
    myLDAfit20 <- LDA(convert(quantdfm, to = "topicmodels"), k = 20)
    get_terms(myLDAfit20, 5)
}
```
