---
title: "Example: Social media analysis (Twitter)"
author: "Kohei Watanabe"
output: 
  html_document:
    toc: true
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "##"
)
```

Using **quanteda**'s `fcm()` and `textplot_network()`, you can perform visual analysis of social media posts in terms of cooccurances of hashtags or usernames in a few steps. The dataset for this example contains only 10,000 Twitter posts, but you can easily analyze more one million posts on your laptop computer.

# Hashtags

```{r, message = FALSE}
library(quanteda)
```

## Load sample data
```{r}
load("data/data_corpus_tweets.RData")
```

## Construct a document-feature matrix of Twitter posts
```{r}
tweet_dfm <- dfm(data_corpus_tweets, remove_punct = TRUE)
head(tweet_dfm)
```

## Extract most common hashtags
```{r}
tag_dfm <- dfm_select(tweet_dfm, ('#*'))
toptag <- names(topfeatures(tag_dfm, 50))
head(toptag)
```

## Construct feature-occurrence matrix of hashtags
```{r}
tag_fcm <- fcm(tag_dfm)
head(tag_fcm)
topgat_fcm <- fcm_select(tag_fcm, toptag)
textplot_network(topgat_fcm, min_freq = 0.1, alpha = 0.8, width = 5)
```

# Usernames

## Extract most frequently mentioned usernames 
```{r}
user_dfm <- dfm_select(tweet_dfm, ('@*'))
topuser <- names(topfeatures(user_dfm, 50))
head(topuser)
```

## Construct feature-occurrence matrix of usernames
```{r}
user_fcm <- fcm(user_dfm)
head(user_fcm)
user_fcm <- fcm_select(user_fcm, topuser)
textplot_network(user_fcm, min_freq = 0.1, col = 'orange', alpha = 0.8, width = 5)
```

